---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(epoxy)
```

# durbanplasticwaste

<!-- badges: start -->

<!-- badges: end -->

## Overview

This packages combines data collected as part of an MSc. Thesis Project and an 
MSc. Semester Project conducted in Durban, South Africa. The projects were 
supported by the Global Health Engineering group at ETH Zurich, Switzerland.

## Installation

You can install the development version of durbanplasticwaste 
from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Global-Health-Engineering/durbanplasticwaste")
```

Alternatively, you can download the individual data sets as a CSV or XLSX file 
from the table below.

```{r, echo=FALSE}
tibble(
  dataset = "litterboom_counts",
  CSV = "[Download CSV](https://github.com/Global-Health-Engineering/durbanplasticwaste/raw/main/inst/extdata/litterboom_counts.csv)",
  XLSX = "[Download XLSX](https://github.com/Global-Health-Engineering/durbanplasticwaste/raw/main/inst/extdata/litterboom_counts.xlsx)"
) |> 
  knitr::kable()
```

## Projects

### MSc. Thesis Project

Evaluating the potential of Extended Producer Responsibility returns for a small
local waste collection company through a brand audit of riverine plastic waste 
in Durban, South Africa.

### Description

This Master's Thesis Project focuses on determining the growth 
opportunities for a small-sized plastic recycling enterprise in light 
of the shift from a voluntary to a mandatory Extended Producer
Responsibility (EPR) policy in South Africa. 

To achieve this goal in  the context of a small start-up in Durban, South Africa
, a brand audit is conducted to identify the top brands that can be targeted for 
financing or partnership opportunities. The company, called 
[TRI ECO Tours](https://triecotours.co.za/), is a small tourism and waste 
collection startup in Durban operated by Siphiwe Rakgabale.

### Research Question

What is the characterization by type, application, and brand of plastic
waste collected in the uMngeni River system in Durban, South Africa?

### Data 

The data was collected throughout two months in Durban, South Africa right 
before the rainy season. The collection took place in 6 different litterboom 
locations throughout Durban. The data gathered was the audit of the occurence 
of the brands washed into the litterbooms. 

The package provides access to three data sets.

```{r}
library(durbanplasticwaste)
```



```{epoxy}
The `litterboom_counts` data set has {ncol(litterboom_counts)} variables and {nrow(litterboom_counts)} observations. For an overview of the variable names, see the following table.  

```

```{r, eval=FALSE}
litterboom_counts
```


```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "litterboom_counts.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

```{epoxy}
The `litterboom_weights` data set has {ncol(litterboom_weights)} variables and {nrow(litterboom_weights)} observations. For an overview of the variable names, see the following table.  

```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "litterboom_weights.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()

```

```{epoxy}
The `locations` data set has {ncol(locations)} variables and {nrow(locations)} observations. For an overview of the variable names, see the following table.  

```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "locations.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()

```

### MSc. Semester Project

<!--# please add semester project title -->

### Description

<!--# please add a short description of the semester -->

### Research Question

<!--# please add the research questions -->

### Data

## Examples

You can find these and more code examples for exploring durbanplasticwaste in 
vignette("examples").

```{r}
library(durbanplasticwaste)
library(dplyr)

litterboom_counts |> 
  group_by(group) |> 
  summarise(
    sum = sum(count)
  ) |> 
  arrange(desc(sum))
```


## License

Data are available as CC-BY.

## Citation

To cite the durbanplasticwaste package, please use:

```{r}
# citation("durbanplasticwaste")
```

