---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(epoxy)
```

# durbanplasticwaste

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/604573987.svg)](https://zenodo.org/badge/latestdoi/604573987)
[![R-CMD-check](https://github.com/Global-Health-Engineering/durbanplasticwaste/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Global-Health-Engineering/durbanplasticwaste/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# Overview

This packages combines data collected as part of an MSc. Thesis Project and an 
MSc. Semester Project conducted in Durban, South Africa. The projects were 
supported by the Global Health Engineering group at ETH Zurich, Switzerland.

# Installation

You can install the development version of durbanplasticwaste 
from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Global-Health-Engineering/durbanplasticwaste")
```

Alternatively, you can download the individual datasets as a CSV or XLSX file 
from the table below.

```{r, echo=FALSE}

extdata_path <- "https://github.com/Global-Health-Engineering/durbanplasticwaste/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  mutate(file_name = str_remove(file_name, ".rda")) |> 
  rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

# Projects

## MSc. Thesis Project

Evaluating the potential of Extended Producer Responsibility returns for a small
local waste collection company through a brand audit of riverine plastic waste 
in Durban, South Africa.

### Description

This Master's Thesis Project focuses on determining the growth 
opportunities for a small-sized plastic recycling enterprise in light 
of the shift from a voluntary to a mandatory Extended Producer
Responsibility (EPR) policy in South Africa. 

To achieve this goal in  the context of a small start-up in Durban, South Africa
, a brand audit is conducted to identify the top brands that can be targeted for 
financing or partnership opportunities. The company, called 
[TRI ECO Tours](https://triecotours.co.za/), is a small tourism and waste 
collection startup in Durban operated by Siphiwe Rakgabale.

### Research Question

What is the characterization by type, application, and brand of plastic
waste collected in the uMngeni River system in Durban, South Africa?

### Data 

The data was collected throughout two months in Durban, South Africa right 
before the rainy season. The collection took place in 6 different litterboom 
locations throughout Durban. The data gathered was the audit of the occurence 
of the brands washed into the litterbooms. 

The package provides access to three data sets.

```{r}
library(durbanplasticwaste)
```



```{epoxy}
The `litterboom_counts` data set has {ncol(litterboom_counts)} variables and {nrow(litterboom_counts)} observations. For an overview of the variable names, see the following table.  

```

```{r, eval=FALSE}
litterboom_counts
```


```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "litterboom_counts.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

```{epoxy}
The `litterboom_weights` data set has {ncol(litterboom_weights)} variables and {nrow(litterboom_weights)} observations. For an overview of the variable names, see the following table.  

```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "litterboom_weights.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()

```

```{epoxy}
The `locations` data set has {ncol(locations)} variables and {nrow(locations)} observations. For an overview of the variable names, see the following table.  

```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "locations.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()

```

## MSc. Semester Project

Examination of non-recycled marine plastic litter in order to identify recycling and beneficiation pathways in Durban, South Africa

### Description

This Semester Thesis Project focuses on determining the distribution of plastic litter on the Durban beachfront in order to identify key targets for policy and financial support through the South African EPR policy to reduce plastic spills into the environment and promote higher recycling rates.
Research Question

What types and amounts of plastic are found along the beachfront in the mangroves of Durban-North, South Africa?

### Data



## Examples

```{epoxy}

The `litterboom_counts` data identifies {length(unique(litterboom_counts$group))} unique groups that own the identified brands. The top 10 brands are shown in the following table. All other brands are lumped together as OTHER. 

```

```{r}
library(durbanplasticwaste)
library(dplyr)
library(forcats)

litterboom_counts |> 
  mutate(group = factor(group)) |> 
  mutate(group = fct_lump(group, n = 10, other_level = "OTHER")) |> 
  group_by(group) |> 
  summarise(
    count = sum(count)
  ) |> 
  arrange(desc(count)) |> 
  mutate(percent = count / sum(count) * 100) |> 
  knitr::kable(digits = 0)

```
  
## License
    
Data are available as [CC-BY](https://github.com/Global-Health-Engineering/durbanplasticwaste/blob/main/LICENSE.md).
    
## Citation
    
To cite the durbanplasticwaste package, please use:

Bergen, R., Schöbitz, L., Meyer-Piening, C., Lin, B., Tilley, E., Kalina, M., Rakgabale, S., Last Name, F. N., & Global Health Engineering. durbanplasticwaste: Durban (South Aftica) Plastic Waste Data (Version 0.1.0) [Data set]. https://doi.org/https://doi.org/10.5281/zenodo.7708629
    

    
